---
title: "lab-07-simpsons.Rmd"
author: "shahad abdulah al- mutairi"
date: "17 March 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages

```{r load_libraries, message=FALSE}
library(tidyverse) 
library(mosaicData)
```


## Exercises


1.  
```{r whickam}
?Whickham
```
    
Your answer:
The data is observational as the description states that is based on age, smoking, and mortality, which are all observable events and not produced via experiments.

2.  
```{r observations}
nrow(Whickham)
```

Your answer;There ane 1,314 observations. As we know every row is an observation.


3.  

```{r number_variables}
names(Whickham)
```

Your answer: There are 3 variables, "outcome" "smoker", and "age"




```{r var_type}
unique(Whickham$outcome)

unique(Whickham$smoker)

unique(Whickham$age)
```

Your answer: Using the `unique()` function on the 3 variables we could see that "outcome" only takes Alive or Dead value, which makes it categorical non-ordinal. "smoker" only takes Yes or No, which also makes it categorical non-ordinal. Age is numerical continous data.

```{r plot_outcome}
ggplot(Whickham, aes(x = outcome)) +

 geom_bar()
```


```{r plot_smoker}
ggplot(Whickham, aes(x = smoker)) +

 geom_bar()
```

```{r plot_age}
ggplot(Whickham, aes(x = age)) +

 geom_histogram()
```


4 Before plotting the correlation between smoking status and health outcomes.I would like to assume there is a strong correlation between the two.based on the well-known impact of smoking on the health.

```{r plot_smokeVSincome_1}
```

Knit, commit, and push to github.

5. 

```{r plot_smokeVSincome_2}

```



6.


```{r age_cat}

```

7. 

```{r plot_smokeVSincome_3}

```

Knit, commit, and push to github.